The \vspec{} Collection: A suite of utilities to model spectroscopic phase curves of 3D exoplanet atmospheres in the presence of stellar variability.

### 2024-08-07 src/tex/ms.tex
Make syw change to allow orchid

@@

% \author[0000-0002-1570-2203]{Ted M Johnson}
\author{Ted M Johnson}
\affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
\affiliation{Center for Research and Exploration in Space Science and Technology, NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
\affiliation{NASA Goddard Space Flight Center \\
8800 Greenbelt Rd \\
Greenbelt, MD 20771, USA}
\affiliation{Nevada Center for Astrophysics, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}

@@

\vspec{} configurations are designed to minimize human errors by providing two equivalent formats. The first, a file written in YAML, is optimized for
human readability. For example, the \texttt{system} section, which describes the relationship between the observed planetary system and the observer, could
be written:
\begin{lstlisting}[
    label={ls:yaml},
    caption={\vspec{} YAML Configuration}
]
system:
    distance: 12.4 pc
    inclination: 89.7 deg
    phase_of_periastron: 0 deg
\end{lstlisting}
Note that the units of each of these parameters is included in a human-readable way. Internally, \vspec{} casts each of these parameters to an Astropy
\texttt{Quantity} instance, so that the user can input any desired value and unit combination so long as the physical type is correct -- i.e. it would be
equivalent to write \texttt{3.83e17 m} as the value for the \texttt{distance} parameter.

The second input method available is to directly initialize the Python object that \vspec{} uses internally to store its model parameters.
The top-level object, \texttt{InternalParameters}, is structured to mirror the YAML input file, with one argument per YAML section. The
parameters in Listing \ref{ls:yaml} would be written:
\begin{lstlisting}[language=Python, caption={\vspec{} Python configuration}]
from astropy import units as u
from VSPEC.params import *
params = InternalParameters(
    ... # Other arguments
    system=SystemParameters(
        distance=12.4*u.pc,
        inclination=89.7*u.deg,
        phase_of_periastron=0*u.deg
    )
)
\end{lstlisting}
This is convenient for producing configurations programmatically, for example for producing a grid of model phase curves.

@@

\section{Interfacing with the Planetary Spectrum Generator via \texttt{libpypsg}}
\label{sec:libpypsg}

PSG\footnote{\url{https://psg.gsfc.nasa.gov/}} \citep{villanueva2018} is a powerful radiative transfer tool that is ubiquitous in the exoplanet and solar system atmosphere fields. Either through its web graphical interface, its public API, or the local version available through Docker, PSG allows users to simulate observations of planets, comets, and moons with a variety of geometries and realistic noise models.

@@

@@

Listing \ref{ls:libpypsg} shows a minimal example that uses \texttt{libpypsg} to make a call to PSG/GlobES. Here \texttt{libpypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance. This configuration object is then passed to an \texttt{APICall} instance. When the API call is made, the contents of the \texttt{PyConfig} are included in the body of the POST request in the \texttt{file} field. PSG reads this the same as it would if we had written those contents to a file \texttt{config.txt} and passed the \texttt{--data-urlencode file@config.txt} option to \texttt{cURL}, as are the instructions in the PSG handbook \citep[pp. 162-163]{villanueva2022}. More details on use of this library including its API and examples can be found on the \texttt{libpypsg} documentation page\footnote{\urldocslibpypsg}.

\begin{lstlisting}[
    basicstyle=\footnotesize,language=Python,label={ls:libpypsg},caption={\texttt{libpypsg} Example}
]
    import libpypsg
    from astropy import units as u
    from netCDF4 import Dataset

    # This next line only needs to be run once per user.
    #     These settings are 
    #     saved to `~/.libpypsg/settings.json`
    libpypsg.settings.save_setting(
        api_key='my_api_key',
        url='https://psg.gsfc.nasa.gov/api.php'
    )

    cfg = libpypsg.cfg.PyConfig(
        target=libpypsg.cfg.Target(
            object='Exoplanet',
            name='proxima-cen-b'
            ...
        ),
        geometry=libpypsg.cfg.Observatory(...),
        generator=libpypsg.cfg.Generator(...),
        telescope=libpypsg.cfg.SingleTelescope(...)
        noise=libpypsg.cfg.CCD(...)
        gcm=libpypsg.globes.waccm_to_pygcm(
            gcm_data,
            itime=0,
            molecules=['H2O','CO2'],
            aerosols=['Water','WaterIce'],
            mean_molecular_mass=48.0
        )

    )

    caller = libpypsg.APICall(
        cfg=cfg,
        app='globes',
        output_type='rad'
    )
    response = caller()

    print(response.rad['Wave/freq','proxima-cen-b'])

    -------------------

    Wave/freq  proxima-cen-b
        um      W / (um m2) 
    ----------- -------------
            1.0   1.81314e-20
           1.02    3.1926e-20
         1.0404   4.95983e-19
       1.061208    5.0892e-19
     1.08243216   3.79632e-19
            ...           ...
     16.3163953   7.66301e-20
     16.6427232   7.66245e-20
    16.97557767   7.54221e-20
    17.31508922   7.49258e-20
    17.66139101   7.35507e-20
    Length = 146 rows

>>>>

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstset{
    backgroundcolor=\color{white},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=1
}

@@

% \author{Ted M Johnson}
\author[0000-0002-1570-2203]{Ted M Johnson}
\affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
\affiliation{Center for Research and Exploration in Space Science and Technology, NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
\affiliation{NASA Goddard Space Flight Center \\
8800 Greenbelt Rd \\
Greenbelt, MD 20771, USA}
\affiliation{Nevada Center for Astrophysics, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}

@@

\vspec{} configurations are designed to minimize human errors by providing two equivalent formats. The first, a file written in YAML, is optimized for
human readability. For example, the \texttt{system} section, which describes the relationship between the observed planetary system and the observer, could
be written as shown in Listing \ref{ls:yaml}.
\begin{lstlisting}[label={ls:yaml},caption=\vspec{} YAML Configuration]
system:
    distance: 12.4 pc
    inclination: 89.7 deg
    phase_of_periastron: 0 deg
\end{lstlisting}
Note that the units of each of these parameters is included in a human-readable way. Internally, \vspec{} casts each of these parameters to an Astropy
\texttt{Quantity} instance, so that the user can input any desired value and unit combination so long as the physical type is correct -- i.e. it would be
equivalent to write \texttt{3.83e17 m} as the value for the \texttt{distance} parameter.

The second input method available is to directly initialize the Python object that \vspec{} uses internally to store its model parameters.
The top-level object, \texttt{InternalParameters}, is structured to mirror the YAML input file, with one argument per YAML section. The same
configuration shown in Listing \ref{ls:yaml} would be written in Python as shown in Listing \ref{ls:python}.
\begin{lstlisting}[language=Python, caption={\vspec{} Python configuration},label={ls:python}]
from astropy import units as u
from VSPEC.params import *
params = InternalParameters(
    ... # Other arguments
    system=SystemParameters(
        distance=12.4*u.pc,
        inclination=89.7*u.deg,
        phase_of_periastron=0*u.deg
    )
)
\end{lstlisting}
This is convenient for producing configurations programmatically, for example for producing a grid of model phase curves.

@@

\section{Interfacing with the Planetary Spectrum Generator via \texttt{libpypsg}}
\label{sec:libpypsg}
PSG\footnote{\url{https://psg.gsfc.nasa.gov/}} \citep{villanueva2018} is a powerful radiative transfer tool that is ubiquitous in the exoplanet and solar system atmosphere fields. Either through its web graphical interface, its public API, or the local version available through Docker, PSG allows users to simulate observations of planets, comets, and moons with a variety of geometries and realistic noise models.

\begin{lstlisting}[
    language=Python,label={ls:libpypsg},
    caption={\texttt{libpypsg} Example}
]
import libpypsg
from astropy import units as u
from netCDF4 import Dataset

# This next line only needs to be run
#     once per user.
#     These settings are 
#     saved to `~/.libpypsg/settings.json`
libpypsg.settings.save_setting(
    api_key='my_api_key',
    url='https://psg.gsfc.nasa.gov/api.php'
)

cfg = libpypsg.cfg.PyConfig(
    target=libpypsg.cfg.Target(
        object='Exoplanet',
        name='proxima-cen-b'
        ...
    ),
    geometry=libpypsg.cfg.Observatory(...),
    generator=libpypsg.cfg.Generator(...),
    telescope=libpypsg.cfg.SingleTelescope(...)
    noise=libpypsg.cfg.CCD(...)
    gcm=libpypsg.globes.waccm_to_pygcm(
        gcm_data, itime=0,
        molecules=['H2O','CO2'],
        aerosols=['Water','WaterIce'],
        mean_molecular_mass=48.0)
)

caller = libpypsg.APICall(
    cfg=cfg, app='globes',
    output_type='rad'
)
response = caller()
print(response.rad['Wave/freq','proxima-cen-b'])
\end{lstlisting}

@@

\begin{lstlisting}[caption={\texttt{libpypsg} Output},label={ls:psgout}]
Wave/freq  proxima-cen-b
    um      W / (um m2) 
----------- -------------
        1.0   1.81314e-20
       1.02    3.1926e-20
     1.0404   4.95983e-19
    ...           ...
16.97557767   7.54221e-20
17.31508922   7.49258e-20
17.66139101   7.35507e-20
Length = 146 rows

@@

Listing \ref{ls:libpypsg} shows a minimal example that uses \texttt{libpypsg} to make a call to PSG/GlobES, with the output shown in Listing \ref{ls:psgout}. Here \texttt{libpypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance. This configuration object is then passed to an \texttt{APICall} instance. When the API call is made, the contents of the \texttt{PyConfig} are included in the body of the POST request in the \texttt{file} field. PSG reads this the same as it would if we had written those contents to a file \texttt{config.txt} and passed the \texttt{--data-urlencode file@config.txt} option to \texttt{cURL}, as are the instructions in the PSG handbook \citep[pp. 162-163]{villanueva2022}. More details on use of this library including its API and examples can be found on the \texttt{libpypsg} documentation page\footnote{\urldocslibpypsg}.
